---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# camtraptor

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/camtraptor)](https://CRAN.R-project.org/package=camtraptor)
[![R-CMD-check](https://github.com/inbo/camtraptor/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/inbo/camtraptor/actions/workflows/R-CMD-check.yaml)
[![codecov](https://codecov.io/gh/inbo/camtraptor/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/camtraptor/)
[![repo status](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
![last commit](https://img.shields.io/github/last-commit/inbo/camtraptor)
<!-- badges: end -->

Camtraptor is an R package to read, explore and visualize Camera Trap Data Packages (Camtrap DP). [Camtrap DP](https://tdwg.github.io/camtrap-dp/) is a community developed data exchange format for this type of data. With camtraptor you can read and filter data, create overviews of observed species, relative abundance or effort, and plot these data on a map.

## Camtrap DP

Camtraptor currently uses the legacy Camtrap DP 0.1.6 for its internal data model. `read_camtrap_dp()` will automatically down-convert Camtrap DP 1.0 datasets to that legacy model to avoid introducing breaking changes. The downside is that some newer properties like `deployments.cameraDepth`, `media.filePublic` and `observations.eventEnd` are removed when reading data.

The upcoming [camtraptor v1.0](https://github.com/inbo/camtraptor/milestone/3) will update the internal data model to Camtrap DP 1.0 and drop support for Camtrap DP 0.1.6. This is a breaking change that will be accompanied by a number of other major changes. Future versions of camtraptor will always use the latest version of Camtrap DP and up-convert legacy datasets to that model.

## Get Started

To get started, see:

- [Vignettes](https://inbo.github.io/camtraptor/articles/): tutorials showcasing functionality.
- [Function reference](https://inbo.github.io/camtraptor/reference/index.html): overview of all functions.

## Installation

You can install the development version of camtraptor from [GitHub](https://github.com/inbo/camtraptor) with:

``` r
# install.packages("devtools")
devtools::install_github("inbo/camtraptor")
```

While we support older versions of R up to 3.5, we recommend using R 4.0.0 or higher.

## Example

Get the taxonomic coverage for an example Camera Trap Data Package `mica`:

```{r}
library(camtraptor)
get_species(mica)
```

Filter observations in `mica` on female mallards and map the number of individuals per deployment location:

```{r}
map_dep(
  mica,
  feature = "n_individuals",
  species = "Anas platyrhynchos",
  sex = "female"
)
```

## camtraptor vs camtrapR and activity

Camtraptor provides and bundles much needed functionality to read, explore and visualize Camera Trap Data Packages. Over time we hope to include some or all of this functionality in widely used R packages such as [camtrapR](https://cran.r-project.org/package=camtrapR) and [activity](https://cran.r-project.org/package=activity).

## Meta

- We welcome [contributions](.github/CONTRIBUTING.md) including bug reports.
- License: MIT
- Get citation information for camtraptor in R doing `citation("camtraptor")`.
- Please note that this project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
