---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# camtraptor

<!-- badges: start -->
[![R-CMD-check-OS](https://github.com/inbo/camtraptor/actions/workflows/check_on_different_r_os.yml/badge.svg)](https://github.com/inbo/camtraptor/actions/workflows/check_on_different_r_os.yml)
[![codecov](https://codecov.io/gh/inbo/camtraptor/branch/main/graph/badge.svg)](https://app.codecov.io/gh/inbo/camtraptor/)
[![repo status](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![funding](https://img.shields.io/static/v1?label=funded+by&message=FWO+for+lifewatch.be&labelColor=1a4e8a&color=f15922)](https://lifewatch.be/)
![last commit](https://img.shields.io/github/last-commit/inbo/camtraptor)
<!-- badges: end -->

Camtraptor is an R package to read, explore and visualize Camera Trap Data Packages ([Camtrap DP](https://tdwg.github.io/camtrap-dp/)). Camtrap DP is a community developed data exchange format for this type of data. With camtraptor you can read and filter data, create overviews of observed species, relative abundance or effort, and plot these data on a map.

To get started, see:

- [Vignettes](https://inbo.github.io/camtraptor/articles/): tutorials showcasing functionality.
- [Function reference](https://inbo.github.io/camtraptor/reference/index.html): overview of all functions.

## Installation

You can install the development version of camtraptor from [GitHub](https://github.com/inbo/camtraptor) with:

``` r
# install.packages("devtools")
devtools::install_github("inbo/camtraptor")
```

## Example

Get the taxonomic coverage for an example Camera Trap Data Package `mica`:

```{r}
library(camtraptor)
get_species(mica)
```

Filter observations in `mica` on female mallards and map the number of individuals per deployment location:

```{r}
map_dep(
  mica,
  feature = "n_individuals",
  species = "Anas platyrhynchos",
  sex = "female"
)
```

## camtraptor vs camtrapR and activity

Camtraptor provides and bundles much needed functionality to read, explore and visualize Camera Trap Data Packages. Over time we hope to include some or all of this functionality in widely used R packages such as [camtrapR](https://cran.r-project.org/package=camtrapR) and [activity](https://cran.r-project.org/package=activity).

## Meta

* We welcome [contributions](.github/CONTRIBUTING.md) including bug reports.
* License: MIT
* Get citation information for `camtraptor` in R doing `citation("camtraptor")`.
* Please note that this project is released with a [Contributor Code of Conduct](.github/CODE_OF_CONDUCT.md). By participating in this project you agree to abide by its terms.
