---
title: "Calculate animal position"
author: "Wolf Missotten"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{calculate-animal-position}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to calculate the **radial distance and angular position** of animals relative to a camera using `calculate_individual_radius_angle()`.

## Setup

Load packages:
```{r load_pkgs}
library(camtraptor)
```

The package includes two built-in example datasets used throughout this vignette: `animal_positions` and `calibration_models`.

`animal_positions` is a data frame containing digitized pixel positions of animals in camera trap images:
```{r show_animal_positions}
animal_positions
```

`calibration_models` is a named list of site calibration models, where each element corresponds to a deployment:
```{r show_calibration_models}
names(calibration_models)
```

## Calculate radius and angle

The function `calculate_individual_radius_angle()` takes the digitized pixel positions and the calibration models and returns the radial distance and angular position of each animal relative to the camera:
```{r basic_usage}
calculate_individual_radius_angle(animal_positions, calibration_models)
```

The function returns the original `animal_positions` data frame with three
additional columns:

column name | description
--- | ---
`radius` | Radial distance from the camera. Units depend on the units used during site calibration.
`angle` | Angular distance from the camera centre line.
`frame_count` | Order of the frame within each sequence (event).

## Specifying column names

By default, the function expects specific column names in `animal_positions` (`deploymentID`, `eventID`, `x`, `y`, `imageWidth`, `imageHeight`). If your data uses different column names, these can be specified explicitly:
```{r custom_columns, eval = FALSE}
calculate_individual_radius_angle(
  animal_positions,
  calibration_models,
  deployment_id = "deploymentID",
  event_id = "eventID",
  x = "x",
  y = "y",
  image_width = "imageWidth",
  image_height = "imageHeight"
)
```