---
title: "Camera operation matrix"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Camera operation matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get a _camera trap station operability matrix_, shortly called _camera operation matrix_ as returned by camtrapR's function [cameraOperation](https://jniedballa.github.io/camtrapR/reference/cameraOperation.html).

## Setup

Load packages:

```{r load_pkgs}
library(camtrapdp)
library(lubridate)
library(dplyr)
```

By loading pacakge `camtrapdp`, a camera trap data package called `camtrapdp` is made available. This data package contains camera trap data of musk rats and coypus. We will use this variable from now on.

## Camera operation matrix

You can create a camera operation matrix by passing a camera trap data package:

```{r basic_usage}
get_cam_op(camtrapdp)
```

Rows are the location IDs (unique values of field `location_name` of `deployments`) preceded by prefix `"Station"` as done by camtrapR's `cameraOperation()`. Column names are dates. The matrix values are: 0 if deployment is not active, 1 or more (integers) if one or more deployments are fully active during all day, a number with a decimal part for start end day as at least one deployment is not fully active durint the entire day.