---
title: "Camera operation matrix"
author: "Damiano Oldoni, Wolf Missotten"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Camera operation matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get a **camera operation matrix** from a Camera Trap Data Package object, equivalent to the matrix returned by camtrapR's function [cameraOperation](https://jniedballa.github.io/camtrapR/reference/cameraOperation.html).

## Setup

Load packages:

```{r load_pkgs}
library(camtraptor)
library(lubridate)
library(dplyr)
```

For this example the function `example_dataset()` is used to load an example Camera Trap Data Package object. The dataset is derived from a study on detecting invasive muskrat and coypu populations using camera traps.

```{r load_dataset}
x <- example_dataset()
```

## Camera operation matrix

The camera operation matrix can be generated with `camtrapR_cameraOperation()`:

```{r basic_usage}
cam_op <- camtrapR_cameraOperation(x)
```
 
For readability, let's only print the first 45 columns of the camera operation matrix (i.e. the first 45 days) instead of all 324:

```{r basic_usage_visualize}
cam_op[, 1:45]
```

To build this matrix, the function reads the `deployments` slot of the Camera Trap Data Package. Row names are station names, by default taken from the `locationName` column of `deployments`. Column names are dates. The matrix values represent the daily effort:

- `NA`: the camera was not set up on that day.
- `1`: the camera was fully active for the entire day.
- A decimal between 0 and 1: the camera was partially active (e.g. on the first or last day of a deployment).
- Greater than `1`: multiple deployments at the same location were active on that day.

In the example above, no location has overlapping deployments, so no values greater than 1 occur in the matrix. To demonstrate the scenario where values are greater than `1`, the example below assigns all four deployments to the same location (`B_HS_val 2_processiepark`):

```{r four_deploys_same_start_end_for_one_location}
x1 <- x
# Assigning all four deployments to the first location
deployments(x1)$locationName[] <- deployments(x1)$locationName[1]
deployments(x1)$deploymentStart[] <- deployments(x1)$deploymentStart[1]
deployments(x1)$deploymentEnd[] <- deployments(x1)$deploymentEnd[1]
# Visualize camera operation matrix
camtrapR_cameraOperation(x1)
```
To demonstrate the scenario where a single location is linked to multiple (non-overlapping) deployments, the example below assigns location `B_HS_val 2_processiepark` to two separate deployments:

```{r two_deploys_two_periods}
x2 <- x
# Assigning the first location to two separate deployments
deployments(x2)$locationName[2] <- deployments(x2)$locationName[1]
deployments(x2)$deploymentStart[2] <- deployments(x2)$deploymentEnd[1] + 
  ddays(5)
deployments(x2)$deploymentEnd[2] <- deployments(x2)$deploymentStart[2] + 
  ddays(5)
# Visualize deployments
deployments(x2) %>% select(locationName, deploymentStart, deploymentEnd)

# Visualize camera operation matrix
camtrapR_cameraOperation(x2)[1, 1:45]
```

### Station names

By default, row names are taken from the `locationName` column of `deployments`. You can specify a different column of `deployments` using `station_col`. The example below uses `locationID`:

```{r cam_op_with_locationID}
cam_op_with_locationID <- camtrapR_cameraOperation(
  x,
  station_col = "locationID"
)
# Since the full matrix would be too wide to display here, we only inspect
# the row names:
row.names(cam_op_with_locationID)
```

### Session and camera IDs

Let's first extend the dataset so `deployments` contains a column for `cameraID` and `sessionID`.

```{r extend_dataset}
x_extended <- x

deployments(x_extended) <- deployments(x) %>% 
  mutate(sessionID = c(1, 2, 3, 4)) %>% 
  mutate(cameraID = c(1, 2, 3, 4))
```

You can specify the column containing the session IDs to be added to the station names using the `session_col` argument. The sessionID will be displayed following camtrapR's convention: `Station__SESS_sessionID`:

```{r add_session_IDs}
cam_op_with_session_ids <- camtrapR_cameraOperation(
  x_extended,
  session_col = "sessionID"
)
# Since the full matrix would be too wide to display here, we only inspect
# the row names:
row.names(cam_op_with_session_ids)
```

You can also specify the column containing the camera IDs to be added to the station names following camtrapR's convention: `Station__CAM_CameraID` using the `camera_col` argument:

```{r add_camera_IDs}
cam_op_with_camera_ids <- camtrapR_cameraOperation(
  x_extended,
  camera_col = "cameraID"
)
# Since the full matrix would be too wide to display here, we only inspect
# the row names:
row.names(cam_op_with_camera_ids)
```

To use both camera and session IDs, the camtrapR's convention `Station__SESS_SessionID__CAM_CameraID` is followed:

```{r add_session_camera_IDs}
cam_op_with_session_and_camera_ids <- camtrapR_cameraOperation(
  x_extended,
  camera_col = "cameraID",
  session_col = "sessionID"
)
# Since the full matrix would be too wide to display here, we only inspect
# the row names:
row.names(cam_op_with_session_and_camera_ids)
```

You can also use the prefix `"Station"` in the station names as done by camtrapR's `cameraOperation()` by setting `use_prefix = TRUE`:

```{r use_prefix}
cam_op_with_session_and_camera_ids_prefix <- camtrapR_cameraOperation(
  x_extended,
  camera_col = "cameraID",
  session_col = "sessionID",
  use_prefix = TRUE
)
# Since the full matrix would be too wide to display here, we only inspect
# the row names:
row.names(cam_op_with_session_and_camera_ids_prefix)
```