---
title: "Camera operation matrix"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Camera operation matrix}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get a _camera trap station operability matrix_, shortly called _camera operation matrix_ as returned by camtrapR's function [cameraOperation](https://jniedballa.github.io/camtrapR/reference/cameraOperation.html).

## Setup

Load packages:

```{r load_pkgs}
library(camtrapdp)
library(lubridate)
library(dplyr)
```

By loading pacakge `camtrapdp`, a camera trap data package called `mica` is made available. This data package contains camera trap data of musk rats and coypus. We will use this variable from now on.

## Camera operation matrix

You can create a camera operation matrix by passing a camera trap data package:

```{r basic_usage}
get_cam_op(mica)
```

The function will read the `deployments` slot of the data package to create the matrix.

The matrix rows are the location IDs (unique values of field `locationName` of `deployments`) preceded by prefix `"Station"` as done by camtrapR's `cameraOperation()`. Column names are dates. The matrix values are: 0 if deployment is not active, an integer (1 or more) if one or more deployments are fully active during all day, a decimal number if at least one deployment is not fully active during the entire day.

In the example above there was a one-to-one relation between deployments and locations: the daily effort was not more than one.

In the example below we show what happens if all four deployments are set to the same location (`B_DL_val 5_beek kleine vijver`):

```{r four_deploys_for_location}
mica1 <- mica
mica1$deployments$locationName <- mica1$deployments$locationName[1]
get_cam_op(mica1)
```

In the example below we simulate a location (`B_DM_val 4_'t WAD`) linked to two deployments active in two different periods:

```{r two_deploys_two_periods}
mica2 <- mica
mica2$deployments$locationName[4] <- mica2$deployments$locationName[3]
mica2$deployments$start[4] <- mica2$deployments$end[3] + ddays(5)
mica2$deployments$end[4] <- mica2$deployments$start[4] + ddays(5)
mica2$deployments %>% select(locationName, start, end)
```

```{r camOp_two_deploys_two_periods}
get_cam_op(mica2)
```
