---
title: "Filter predicates"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Filter predicates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows what filter predicates are and how to use them in `get_*()` functions or in `map_dep()`.

## Setup

Load packages:

```{r load_pkgs}
library(camtrapdp)
library(lubridate)
library(dplyr)
```

By loading pacakge `camtrapdp`, a camera trap data package called `camtrapdp` is made available. This data package contains camera trap data of musk rats and coypus. We will use this variable from now on.

## Filter predicates

All filter predicates are functions starting with `pred` prefix. They can be distinguished in four categories based on the type of inputs they accept:

1. one argument, one value
2. one argument, no value
3. one argument, multiple values (vector)
4. multiple predicates

They are called filter predicates because they build (dplyr) filter statement. Filter predicates return objects of class  `filter_predicate`, which are a particular kind of list with the following slots:

1. `arg`, the argument
2. `value`, the value
3. `type`, the type of filter predicate
4. `expr`, the filter dplyr expression

### One argument - one value predicates

This filter predicates accept one argument and one value.

#### `pred()`, `pred_not()`

The `pred()` is the most basic predicates and refers to equality statements. Example, if you want to select rows where column `a` is equal to 5:

```{r pred}
pred(arg = "a", value = 5)
```

The opposite operator of `pred()` (equals) is `pred_not` (notEquals):

```{r pred_not}
pred_not(arg = "a", value = 5)
```

#### `pred_gt()`, `pred_gte()`, `pred_lt()`, `pred_lte()`

These predicates express `>` (greaterThan), `>=` (greaterThanOrEqual),`<` (lessThan) and `<=` (lessThanOrEqual) respectively. Example: if you want to select rows where column `a` is greater than 5:

```{r pred_gt}
pred_gt(arg = "a", value = 5)
```

### One argument - no value predicates

The predicate `pred_na()` compares the argument against NA. To select rows where column `a` is NA: 

```{r pred_na}
pred_na(arg = "a")
```

To select all the rows where `a` is not NA, you can use the opposite predicate `pred_notna()`:

```{r pred_notna}
pred_notna(arg = "a")
```

### One argument - multiple value predicates

These predicates accept a vector with multiple values as argument. To get rows where column `a` is one of the values `c(1,3,5)`:

```{r pred_in}
pred_in(arg = "a", value = c(1,3,5))
```

The opposite of `pred_in()` is  `pred_notin()`:

```{r pred_notin}
pred_notin(arg = "a", value = c(1,3,5))
```

### multiple predicates: `pred_and()` and  `pred_or()`

You can combine the predicates described above to make more complex filter statements by using `pred_and()` (AND operator) and  `pred_or()` (OR operator).

Some examples. Select rows where column  `a` is equal to 5 and column `b` is not NA:

```{r pred_and}
pred_and(pred("a", 5), pred_notna("b"))
```

Select rows where column `a` is equal to 5 or column `b` is not NA:

```{r pred_or}
pred_or(pred("a", 5), pred_notna("b"))
```

Notice how these two predicates return `filter_predicate` objects with the same structure as any other predicate, but with slots  `arg`, `value` and  `type` as long as the number of input predicates they combine.

## How to use filter predicates

The filter predicates are useful to select a subset of deployments for the `get_*()` functions and the visualization function `map_dep()`.

### One predicate

Apply get_* functions only  to the deployments with location name _B_ML_val 06_Oostpolderkreek_ or _ _B_ML_val 07_Sint-Anna_:

```{r example_location_name_get_n_obs}
get_n_obs(camtrapdp,
          pred_in("location_name",
                  c("B_ML_val 06_Oostpolderkreek", "B_ML_val 07_Sint-Anna")))
```

```{r example_location_name_get_effort}
get_effort(camtrapdp,
           pred_in("location_name",
                   c("B_ML_val 06_Oostpolderkreek", "B_ML_val 07_Sint-Anna")))
```

```{r example_location_name_get_n_species}
get_n_species(camtrapdp,
              pred_in("location_name",
                      c("B_ML_val 06_Oostpolderkreek", "B_ML_val 07_Sint-Anna")))
```

### Multiple predicates

As shown above, you can combine several predicates for more complex filtering. E.g. calculate the number of species detected by the deployments with one of the location names _B_ML_val 06_Oostpolderkreek_ and _B_ML_val 07_Sint-Anna_, or deployments souther than 51.28 degrees:

```{r pred_or_get_n_species}
get_n_species(camtrapdp,
              pred_or(
                pred_in("location_name",
                        c("B_ML_val 06_Oostpolderkreek", "B_ML_val 07_Sint-Anna")),
                pred_lt("latitude", 51.28)))
```

Same syntax is valid for visualizing such information via `map_dep()` function:

```{r exmaple_map_Dep_with_multiple_predicates}
map_dep(camtrapdp,
        feature = "n_species", 
        pred_or(
          pred_in("location_name",
                  c("B_ML_val 06_Oostpolderkreek", "B_ML_val 07_Sint-Anna")),
          pred_lt("latitude", 51.28)))
```

Notice also that you can pass as much predicates as you want to get_* functions or `map_dep()` by separating them with comma: they will be combined internally using the AND operator. E.g. to get effort of deployments southern than 51.28 degrees AND eastern than 3.59 degrees, you can simplify this:

```{r example_pred_and_get_n_species}
get_n_species(camtrapdp,
              pred_and(pred_lt("latitude", 51.28),
                       pred_gt("longitude", 3.59)))
```

by omitting the  `pred_and()`:

```{r example_pred_and_get_n_species_simplified}
get_n_species(camtrapdp,
              pred_lt("latitude", 51.28),
              pred_gt("longitude", 3.59))
```

This is similar to the behvaior of dplyr's  `filter()` function, where this:

```{r example_dplyr_filter_and}
camtrapdp$deployments %>%
  filter(latitude < 51.28 & longitude > 3.59)
```

is exactly the same as this:

```{r example_dplyr_filter_and_simplified}
camtrapdp$deployments %>%
  filter(latitude < 51.28, longitude > 3.59)
```

Happy filtering!
