---
title: "Record table"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Record table}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to get a _species record table_ as returned by camtrapR's function [recordTable](https://jniedballa.github.io/camtrapR/reference/recordTable.html) starting from a Camera Trap Data Package object.

## Setup

Load packages:

```{r load_pkgs}
library(camtraptor)
library(lubridate)
```

For this example the function `example_dataset()` is used to load an example Camera Trap Data Package object. The dataset is derived from a study on detecting invasive muskrat and coypu populations in Belgium.

```{r load_dataset}
x <- example_dataset()
```

## Species record table

The camtrapR function `recordTable()` generates a record table from camera trap images or videos. In a Camera Trap Data Package, `media` (e.g. images) are already aggregated into `observations` to a certain extent. If all observations are considered independent, the record table can be generated simply with:

```{r basic_usage}
camtrapR_recordTable(x)
```

The function returns the same columns as the camtrapR's function `recordTable()` except for column `n`. The following mapping is applied:

column name output | description
--- | ---
`Station` | the station name as provided by argument `stationCol` (default: `locationName`). It has to be a column of `deployments`
`Species` |  the `scientificName` column in `observations`
`n` | the `count` column in `observations` (number of observed individuals)
`DateTimeOriginal` | the `eventStart` column in `observations`
`Date` | the date from `eventStart`
`Time` | the time part from `eventStart`
`delta.time.secs` | the elapsed time in seconds between two (independent) observations
`delta.time.mins` | the elapsed time in minutes between two (independent) observations
`delta.time.hours` | the elapsed time in hours between two (independent) observations
`delta.time.days` | the elapsed time in days between two (independent) observations
`Directory` | a list with file paths as stored in column `filePath` of `media`
`FileName` | a list with file names as stored in column `fileName` of `media`
`Latitude` | the latitude of the station
`Longitude` | the longitude of the station
`clock` | the clock time in radians
`solar` | the solar time in radians

The following remarks are both valid for camtrapR's function  `recordTable()` and the function  `camtrapR_recordTable()` of this package:

1. observations are grouped by station and species
2. observations of unidentified animals are removed
2. the elapsed time of the first observation (record) of a species at a certain station is set to 0 by default

### Temporal independence

As described in [Chapter 3](https://jniedballa.github.io/camtrapR/articles/camtrapr3.html) of camtrapR documentation, observations can be filtered for temporal independence by setting a minimum time difference between subsequent records of the same species. As for `recordTable()`, this is achieved via argument `minDeltaTime`, defined as the minimum time difference (in minutes) between two records of the same species at the same station which are to be considered independent. The default is 0, causing the function to return all records.

Again, as for `recordTable()`, an argument is provided, `deltaTimeComparedTo`, to further control how independence between records is assessed. 

For each event there are generally multiple media files, as many camera traps use bursts. Setting `deltaTimeComparedTo` to `"lastIndependentRecord"` returns only records taken `minDeltaTime` minutes after the first photo of the last independent record (i.e., `eventStart`). Setting `deltaTimeComparedTo` to `"lastRecord"` returns only records taken `minDeltaTime` minutes after the last photo of the last independent record (i.e., the last `timestamp` in the event's media).


Let's first modify the temporal data of the Camera Trap Data Package object to create a dependent observation, solely for demonstration purposes.

```{r dummy_data_example}
obs <- observations(x)
obs[obs$eventID == "02ae9f43", "eventStart"] <- as_datetime("2020-08-02 05:10:20")
 
med <- media(x) 
rows_to_update <- which(med$eventID == "02ae9f43") 
med[rows_to_update, "timestamp"] <- as_datetime("2020-08-02 05:10:20") 

x_modified <- x
observations(x_modified) <- obs
media(x_modified) <- med
```

Example with `minDeltaTime = 10` and `deltaTimeComparedTo = "lastIndependentRecord"`:

```{r example_lastIndependetRecord}
camtrapR_recordTable(
  x_modified,
  minDeltaTime = 10,
  deltaTimeComparedTo = "lastIndependentRecord"
  )
```

Example with `minDeltaTime = 10` and `deltaTimeComparedTo = "lastRecord"`:

```{r example_lastRecord}
camtrapR_recordTable(
  x_modified,
  minDeltaTime = 10,
  deltaTimeComparedTo = "lastRecord"
  )
```
Running the code above with `deltaTimeComparedTo = "lastRecord"` produces the message `Number of not independent observations to be removed: 1`, which does not appear when using `deltaTimeComparedTo = "lastIndependentRecord"`. The two settings thus produce different record tables, compare row 5 in particular.

This difference arises because event `45ee3031` spans a 51-second burst. `lastRecord` counts from the last photo of that burst (05:01:05), whereas `lastIndependentRecord` counts from the first photo (05:00:14). The next Anas platyrhynchos event (05:10:20) falls between the two thresholds: it is independent under lastIndependentRecord (threshold: 05:10:14) but dependent under lastRecord (threshold: 05:11:05).

### Exclude some species

Similar to `recordTable()`, the function `camtrapR_recordTable()` allows you also to exclude species. Only scientific names are allowed:

```{r example_exclude}
camtrapR_recordTable(x, exclude = c("Anas platyrhynchos", "Vulpes vulpes"))
```

### Station names

The column containing the station names can also be defined by the user if the default value, `"locationName"`, is not the correct one. It has to be a valid column of  `deployments`. Here below, `locationID` is used:

```{r example_stationCol}
camtrapR_recordTable(x, stationCol = "locationID")
```

### Remove duplicates

Sometimes multiple observations of the same species are recorded at the same time and location, but differ in attributes such as `lifeStage` or `sex`, for example, when adult females and juveniles are photographed together. The `removeDuplicateRecords` argument controls whether such duplicates are retained: by default it is `TRUE`, keeping only one observation per group. Set it to `FALSE` to retain all observations.

To illustrate the difference, compare the two record tables below.

Without duplicates (default):

```{r no_duplicates}
camtrapR_recordTable(x)
```

With duplicates:

```{r duplicates_allowed}
camtrapR_recordTable(x, removeDuplicateRecords = FALSE)
```

Row 10 differs between the two outputs (among others). When `removeDuplicateRecords = FALSE`, there are two records with`DateTimeOriginal = 2020-06-06 04:11:07` one on row 9 and one on row 10.

### Other arguments needed?

Are there other arguments of camtrapR's function `recordTable()` you think should be relevant to add to `camtrapR_recordTable()`, please let us know by posting an [issue](https://github.com/inbo/camtraptor/issues)!
