---
title: "Visualize number of species per deployment"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualize number of species per deployment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows how to visualize number of species per deployment in a leaflet by using function  `visualize_species_per_deployment()`.

## Read data

First, import camera trap data. This package accepts only data formatted using the TDWG [Camera Trap Data Pakckage](https://tdwg.github.io/camtrap-dp/) standard.

To read data, use the function `read_camtrap-dp()`:

```{r read_data, message=FALSE, warning=FALSE}
library(here)
camtrap_dp_folder <- here("data", "gmu8-monitoring-faunabeheerzone-8-20210301093537")
camtrap_dp <- read_camtrap_dp(camtrap_dp_folder)
```

## Create map

You can visualize the number of species detected by each deployment by using the function `visualize_species_per_deployment()`. 

You can benefit of the `cluster` mode (set to `TRUE` by default):

```{r visualize_with_clusters}
visualize_species_per_deployment(deployments = camtrap_dp$deployments,
                                 observations = camtrap_dp$observations)
```

or you can deactivate it:

```{r visualize_without_clusters}
visualize_species_per_deployment(deployments = camtrap_dp$deployments,
                                 observations = camtrap_dp$observations,
                                 cluster = FALSE)
```



Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
