---
title: "Visualize deployment features"
author: "Damiano Oldoni, Wolf Missotten"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualize deployment features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette shows you how to use the function `map_summary()` to visualize important features for each deployment on an interactive leaflet map such as:

-   number of detected species
-   number of events observed
-   number of observations
-   number of individuals observed
-   RAI (Relative Abundance Index) (calculated based on observations or individuals)
-   effort (duration of a deployment)

## Setup

Load the packages that will be used in this example:

```{r load_libraries}
library(camtraptor)
library(dplyr)
```

For this example the function `example_dataset()` is used to load an example Camera Trap Data Package dataset. The dataset is derived from a study on detecting invasive muskrat and coypu populations using camera traps.

```{r load_dataset}
x <- example_dataset()
```

Before making a visualisation with `map_summary()`, the Camera Trap Data Package dataset should be transformed to a grouped data frame. This can be done by using the function `summarize_observations()` or `summarize_deployments()`, depending on what aspect of the dataset you want to visualize. Both are demonstrated in this example.

```{r summarize_to_df}
summ_obs <- x %>%
  summarize_observations(group_by = c("deploymentID", "latitude", "longitude"))

summ_depl <- x %>%
  summarize_deployments(group_by = c("deploymentID", "latitude", "longitude"))
```

Let's also create a Camera Trap Data Package dataset that is filtered for a specific species.

To do so, you can first inspect which species have been detected. The function `taxa()` gives you an overview of the unique scientific names within the Camera Trap Data Package dataset:

```{r species_list}
taxa(x)
```

This example filters for Anas platyrhynchos. You can filter by using the `filter_observations()` function:

```{r filter_observations}
x_anas_p <- x %>% 
  filter_observations(scientificName == "Anas platyrhynchos")

summ_obs_anas_p <- x_anas_p %>%
  summarize_observations(group_by = c("deploymentID", "latitude", "longitude"))
```

## Create maps

### Basic usage

#### Number of detected species

You can visualize the number of detected species by each deployment by using the function `map_summary()` with `feature` set to `n_scientificName`:

```{r visualize_n_species}
 summ_obs %>%
  map_summary(feature = "n_scientificName")
```

#### Number of events observed

To visualize the number of events observed, set `feature = n_events`:

```{r visualize_n_events}
summ_obs %>%
  map_summary(feature = "n_events")
```

#### Number of observations

To visualize the number of observations, set `feature = n_observations`:

```{r visualize_n_obs}
summ_obs %>%
  map_summary(feature = "n_observations")
```

You can also visualize the number of observations for a specific species: 

```{r visualize_n_obs_species}
summ_obs_anas_p %>%
  map_summary(feature = "n_observations")
```

When the argument `extend` is set as `TRUE` within the `summarize_observations()` function, the deployments without observations for Anas platyrhynchos will also be visualized by a black multiplication symbol `×`:

```{r visualize_n_obs_species_with_extend}
# Make grouped data frame by summarizing
summ_obs_anas_p_extend <- x_anas_p %>%
  summarize_observations(
    group_by = c("deploymentID", "latitude", "longitude"), 
    extend = TRUE
  )

# Visualize
summ_obs_anas_p_extend %>%
  map_summary(feature = "n_observations")
```

#### Number of individuals

To visualize the number of observed individuals, set `feature = sum_count`:

```{r visualize_n_individuals_with_clusters_and_default_hover_col}
summ_obs %>%
  map_summary(feature = "sum_count")
```

As for observations, you can also visualize the number of observed individuals for a specific species: 

```{r specify_species_individuals}
summ_obs_anas_p %>%
  map_summary(feature = "sum_count")
```

#### RAI (observations)

> **Note:** Notice that in this package the RAI is normalized over a deployment activity period of 100 days.

To visualize the Relative Abundance Index (RAI) for a given species (e.g. Anas platyrhynchos), set `feature = "rai_observations"`. This RAI is based on the number of observations, see the next section on how to calculate the RAI on the number of detected individuals.

```{r show_rai_observations}
summ_obs_anas_p %>%
  map_summary(feature = "rai_observations")
```

#### RAI (individuals)

> **Note:** Notice that in this package the RAI is normalized over a deployment activity period of 100 days.

You can also visualize the RAI based on the number of detected individuals instead of the number of observations (see previous section). Set `feature = "rai_count"`:

```{r show_rai_individuals}
summ_obs_anas_p %>%
  map_summary(feature = "rai_count")
```

#### Effort

You can visualize the duration of the deployments, also called *effort*, as number of active hours by using the grouped data frame created by `summarize_deployments()` in the section 'Setup'. Set `feature = "effort_duration"`:

```{r effort_in_hours}
summ_depl %>%
  map_summary(feature = "effort_duration")
```

Other units than 'hour' (which is the default) can also be used, by using the argument `effort_unit`. For example, you can express the effort in days:

```{r effort_in_days}
summ_depl %>%
  map_summary(feature = "effort_duration", effort_unit = "day")
```

### Visualize deployments without detected animals

It can happen that some deployments didn't detected any recognizable animal (`scientificName` = `NA`) or didn't observe anything at all (deployments with no observations). While visualizing the number of species, these two situations are shown by default as black and red multiplication symbols `×` respectively:

```{r show_deployments_without_observations_or_without_known_species}
# Create a data package demonstrating two scenarios:
# 1. Deployments with observations of unknown species (black ×)
# 2. Deployments with no observations at all (red ×)

# Keep only observations with unknown species
x_unknown <- x %>%
  filter_observations(is.na(scientificName))

# Remove all observations from deployment "00a2c20d"
# to simulate a deployment with no observations at all
x_unknown <- x_unknown %>%
  filter_observations(deploymentID != "00a2c20d")

# Summarize, extend = TRUE ensures deployments with no observations are included
summary_obs_unknown <- x_unknown %>%
  summarize_observations(
    group_by = c("deploymentID", "latitude", "longitude"),
    extend = TRUE
  )

# Visualize
summary_obs_unknown %>%
  map_summary(feature = "n_scientificName")
```

To hide deployments with unknown species or deployments with no observations, set `na_values_show` or `zero_values_show` to `FALSE` respectively.

### Clustering and hovering

You can control which variables are displayed when hovering over a deployment using the `hover_columns` argument. You can choose among all columns from the grouped data frame created with `summarize_observations()` or `summarize_deployments()`.

In the example below only the deploymentID and the number of observations are shown while hovering:

```{r specify_hover_col}
summ_obs %>%
  map_summary(
    feature  = "n_observations",
    hover_columns = c("deploymentID", "n_observations")
  )
```

Deactivating both cluster mode and hovering is also possible:

```{r visualize_without_clusters_and_hovering}
summ_obs %>% 
  map_summary(feature = "n_observations", cluster = FALSE, hover_columns = NULL)
```

### Styling

#### Use a color palette

The default color palette is a [viridis color palette](https://cran.microsoft.com/snapshot/2017-08-01/web/packages/viridis/vignettes/intro-to-viridis.html) called `"inferno"`. You can specify another viridis color palette, e.g. `"viridis"` or `"magma"`, or a [RColorBrewer](https://renenyffenegger.ch/notes/development/languages/R/packages/RColorBrewer/index) palette, e.g. `"BuPu"` or `"Oranges"`. Below the `viridis` color palette is used:

```{r use_viridis}
summ_obs %>% 
  map_summary(feature = "n_observations", palette = "viridis")
```

You can use a palette from `RColorBrewer`, e.g. the `"BuPu"` palette:

```{r use_BuPu}
summ_obs %>% 
  map_summary(feature = "n_observations", palette = "BuPu")
```

Another easy way to specify a palette is to create it by passing a vector of colors as names or hex colors, e.g. `c("black", "blue", "#A3675F")`:

```{r use_black_blue_red}
summ_obs %>% 
  map_summary(
    feature = "n_observations",
    palette = c("black", "blue", "#A3675F")
  )
```

#### Use a specific icon and color for zero values

You can customise the icon shown for zero-value deployments by passing an icon URL to `zero_values_icon_url` and its size in pixels to `zero_values_icon_size`. Many free icon libraries are available online, such as [icons8](https://icons8.com/). The following example uses a custom icon with a size of 50 pixels:

```{r specify_zero_icon}
x_anas_p %>%
  summarize_observations(
    group_by = c("deploymentID", "latitude", "longitude"),
    extend = TRUE
  ) %>%
  map_summary(
    feature = "n_observations",
    zero_values_icon_url = "https://img.icons8.com/color/48/000000/futurama-fry.png",
    zero_values_icon_size = 50
  )
```

Typically the colour is part of the URL. Here below is an example where you change the color of the default icon to green (2ECC71):

```{r specify_zero_colors_value}
x_anas_p %>%
  summarize_observations(
    group_by = c("deploymentID", "latitude", "longitude"),
    extend = TRUE
  ) %>%
  map_summary(
    feature = "n_observations",
    zero_values_icon_url = "https://img.icons8.com/ios-glyphs/30/2ECC71/multiply.png"
  )
```

Modifying the default value (`"black"`) can be useful as the color of deployments with zero values can be sometimes too similar to one of the colors used in the palette.

#### Modify circle size

You can also modify the upper and lower limit of the circle sizes by specifying `radius_range` (default:`c(10,50)`):

```{r change_size_circles}
summ_obs %>% 
  map_summary(feature = "n_observations", radius_range = c(20, 150))
```

#### Use absolute scale

By default the upper limit of the color palette and radius are defined based on the actual feature values. However, sometimes it can be useful to set up an absolute upper limit. This can be done by setting argument `relative_scale` to `FALSE` and specifying the upper limit in `max_scale`.

Upper limit lower than number of observations:

```{r low_upper_limit}
summ_obs %>% 
  map_summary(
    feature = "n_observations",
    relative_scale = FALSE,
    max_scale = 2
  )
```

Upper limit higher than number of observations:

```{r high_upper_limit}
summ_obs %>% 
  map_summary(
    feature = "n_observations",
    relative_scale = FALSE,
    max_scale = 50
  )
```
