---
title: "Workflow of the package"
author: "Wolf Missotten"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Workflow of the package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Camtraptor is an R package to **read**, **explore** and  **visualize** [Camera Trap Data Package](https://camtrap-dp.tdwg.org/) (Camtrap DP) datasets. This vignette walks you through the workflow of the package.

> **Note:** Coming from `camtraptor` version 0.xx? Check out [News](https://inbo.github.io/camtraptor/news/index.html) to see what changed. Deprecated functions will throw a warning pointing you towards their replacements.

## Setup

Load the package:

```{r setup}
library(camtraptor)
```

## Workflow

### Reading data

To start your data analysis you of course first need data.

Here, the function `example_dataset()` is used to load an example Camera Trap Data Package dataset that is included in the camtraptor package. This dataset is derived from a study on detecting invasive muskrat and coypu populations using camera traps.

```{r load_dataset}
x <- example_dataset()
```

To read your own locally stored dataset, use `read_camtrapdp()`:

```{r read_own_data, eval = FALSE}
x <- read_camtrapdp("path/to/datapackage.json")
```
In this vignette the example dataset will be used.

### Exploring data

Now that you read in your data, you can start exploring it. A Camtrap DP dataset consists of three tables: `deployments`, `media` and `observations`. For more details on the data structure, see the [Camtrap DP website](https://camtrap-dp.tdwg.org/data/). You can access each table directly by using `deployments()`, `media()` and `observations()`. Let's take a look at the deployments table: 

```{r deployments}
deployments(x)
```
To get a quick overview of which species were recorded, use `taxa()`:

```{r taxa}
taxa(x)
```

### Filtering data

Now that you have explored the dataset, you can start filtering. Use `filter_deployments()` to select specific deployment locations and `filter_observations()` to select specific species. For example, filtering for two specific locations and mallard (*Anas platyrhynchos*):

```{r filter_mallard_location}
x_filtered <- x %>%
  filter_deployments(
    locationName == c(
      "B_HS_val 2_processiepark",
      "B_DM_val 4_'t WAD"
      )
    ) %>%
  filter_observations(scientificName == "Anas platyrhynchos")
```

`filter_deployments()` filters the deployments table to the two selected locations, and automatically removes any associated observations and media that do not belong to these deployments. `filter_observations()` then filters the remaining observations to only keep mallard (*Anas platyrhynchos*) observations.

> **Note:** Want to filter out timelapse observations? Use `filter_out_timelapse()` as a shortcut for `filter_observations(x, captureMethod != "timelapse")`.


### Summarizing data

Before visualizing, the data first needs to be summarized.

Use `summarize_observations()` to get an overview of the observations per deployment:

```{r summarize_observations}
summ_obs_filtered <- summarize_observations(x_filtered)
summ_obs_filtered
```

Use `summarize_deployments()` to get an overview of the deployments:

```{r summarize_deployments}
summ_depl_filtered <- summarize_deployments(x_filtered)
summ_depl_filtered
```

### Visualizing data

Use `map_summary()` to visualize a summary on a map.

Visualize the number of mallard observations per deployment:

```{r map_observations}
map_summary(summ_obs_filtered, feature = "n_observations")
```
As you can see, in only one of the two deployment was mallard observed.


You can also visualize the daily effort per deployment:

```{r map_effort}
map_summary(summ_depl_filtered, feature = "effort_duration")
```

### Putting it all together

In practice you'd chain everything into a single pipeline, e.g.:

```{r full_pipeline}
x %>%
  filter_observations(scientificName == "Anas platyrhynchos") %>%
  filter_deployments(
    locationName == c(
      "B_HS_val 2_processiepark",
      "B_DM_val 4_'t WAD"
      )
    ) %>%
  summarize_observations() %>%
  map_summary(feature = "n_observations")
```

  > **Note**: For more information about visualization, checkout [the next vignette about visualizing](visualize-deployments-features.html)